On centrality of orthogonal $\Kt$

### 2016-04-27 paper.tex
Improved proof of the isomorphism St*=St

The Steinberg group $\St(n, R) = \St(\rA_{n-1}, R)$ is the group presented by generators
$x_{ij}(r)$, $1\leq i\neq j\leq n$, $r\in R\}$ and relations
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\begin{align}
&x_{ij}(r)x_{ij}(s)=x_{ij}(r+s),\\
&[x_{ij}(r),\,x_{hk}(s)]=1,\text{ for }h\neq j,\ k\neq i,\\
&[x_{ij}(r),\,x_{jk}(s)]=x_{ik}(rs).
\end{align}
The natural projection $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ sends $x_{ij}(r)$ to $t_{ij}(r)$.

@@

Recall that an ideal $I\trianglelefteq R$ is called a \emph{splitting ideal} if the canonical projection $R \twoheadrightarrow R/I$ splits as a unital ring morphism.
For a splitting ideal $I\trianglelefteq R$ the \emph{relative Steinberg group} $\St(n,\,R,\,I) = \St(\rA_{n-1}, R, I)$ can be defined as the kernel $\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$.

For a general $I$ the relative Steinberg group $\St(\Phi, R, I)$ is no more a subgroup of $\St(\Phi, R)$ but rather is its central extension.
For the purposes of the present text it suffices to consider only relative Steinberg groups corresponding to splitting ideals $I$.
For more information regarding general relative Steinberg groups we refer the reader to \cite[Section~3]{SCh}.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
%\begin{align} \xymatrix{0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0} \end{align} splits, i.e., $\pi\sigma=1$.
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.

@@

\begin{df}
For $I\trianglelefteq R$ and $n\geq4$ define $\St^*(n,\,R,\,I)$ to be the group with the set of generators
\begin{multline*}
\{F(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}\,\cup\{S(u,\,v)\mid u\in I^n,\ v\in\E(n,\,R)e_1,\ u^tv=0\}
\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)\inv=F(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj3} \\
&F(u,\,va)=S(ua,\,v)\ \forall\,a\in I,\,(u,\,v^t)=(M e_1,\,e_2^t M\inv),\,M\!\in\E(n,\,R) \label{coinc}.
\end{align}
\end{df}
Notice that we only described how a generator $F(u, v)$ acts via conjugation on another generator $F(u',v')$ and omitted three similar relations involving generators $S(u,v)$ of the second type.
The reason for this is the following lemma which asserts that the ``missing'' relations will follow automatically from \eqref{add4}--\eqref{coinc}.
\begin{lm}
\label{allyouneedisf}
Denote by $\phi\colon\St^*(n,\,R,\,I)\rightarrow\E(n,\,R)$ the natural map sending $F(u,\,v)\mapsto t(u,\,v)$ and $S(u,\,v)\mapsto t(u,\,v)$.
Then the following facts are true.
\begin{lmlist}
\item $\St^*(n,\,R,\,I)$ is generated as an abstract group by the set of elements 
      $$\{F(u,\,va)\mid a\in I,\ (u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\in\E(n,\,R)\};$$
\item for any $g\in\St^*(n,\,R,\,I)$ one has 
      \setcounter{equation}{2} \renewcommand{\theequation}{R\arabic{equation}'}
      \begin{equation} gF(u,\,v)g\inv=F(\phi(g)u,\,\phi(g\inv)^tv); \end{equation}
\item for any $g\in\St^*(n,\,R,\,I)$ one has
       \setcounter{equation}{2} \renewcommand{\theequation}{R\arabic{equation}''}
      \begin{equation} gS(u,\,v)g\inv=S(\phi(g)u,\,\phi(g\inv)^tv); \end{equation}
\item there is a ``transpose automorphism'' defined on $\St^*(n, R, I)$ satisfying
      $$F(u,\,v)^t=S(v,\,u),\quad S(u,\,v)^t=F(v,\,u).$$
\end{lmlist} \end{lm}

@@

Firstly, we define $\iota$ as a homomorphism to the absolute group (for which we de not need the assumption that $I$ is splitting).
$$\xymatrix{\St^*(n,\,R,\,I)\ar@<0.0ex>[r]^{\iota}&\St(n,\,R).}$$

Clearly, we must map the elements $F(u,\,v)$ to van der Kallen's generators $X(u,\,v)$.
Now we should find the images for elements $S(u,\,v)$ as well.
First of all, recall that van der Kallen in~\cite[3.8--3.10]{vdK} constructs the following elements
$$x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.$$
The following definition is the ``transpose'' of \cite[3.13]{vdK}.
\begin{df} For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$
 consisiting of all elements $y\in\St(n,\,R)$ that admit the decomposition $y=\prod x(w^k,\,v)$, 
 where $\sum u^k=u$ and $u^k=c^k \cdot (e_pv_q-e_qv_p)$ for some $c^k\in R$, $1\leq p\neq q\leq n$. \end{df}

@@

In order to construct the arrow $\kappa$ we use the presentation from \cref{prop:TulPres}.
We set $$\kappa(X(u,v)) = F(u,v),\ u\in \E(n, R)e_1,\ v\in I^n.$$ 
We have to check that $F(u,v)$ satisfy the relations \eqref{add2}--\eqref{add3}.
The relations \eqref{add2} and \eqref{conj2} are immediately obvious.

@@

At this point we obtained the presentation $\St^*(n,\,R,\,I)$ for the $\St(n,\,R,\,I)$ where $I\trianglelefteq R$ is a splitting ideal, $n\geq4$.

Below we obtain the main result of this section, namely, construct a map 
$$
\mathrm T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X],\,XR_a[X]),
$$
for $n\geq4$. A local-global principle for Steinberg group and centrality of $\mathrm K_2$ formally follow from existance of this map~\cite[Lem.~15, Lem.~16, proof of Th.~2]{SCh}.

>>>>

The Steinberg group $\St(n, R) = \St(\rA_{n-1}, R)$ is the group presented by generators
$x_{ij}(r)$, $1\leq i\neq j\leq n$, $r\in R\}$ and relations
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\begin{align}
&x_{ij}(r)x_{ij}(s)=x_{ij}(r+s), \label{add0}\\
&[x_{ij}(r),\,x_{hk}(s)]=1,\text{ for }h\neq j,\ k\neq i, \label{ccf1}\\
&[x_{ij}(r),\,x_{jk}(s)]=x_{ik}(rs) \label{ccf2}.
\end{align}
The natural projection $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ sends $x_{ij}(r)$ to $t_{ij}(r)$.

@@

Recall that an ideal $I\trianglelefteq R$ is called a \emph{splitting ideal} if the canonical projection $R \twoheadrightarrow R/I$ splits as a unital ring morphism, i.\,e. one has $\pi\sigma=1$.
$$\xymatrix{0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0}$$
For a splitting ideal $I\trianglelefteq R$ the \emph{relative Steinberg group} $\St(n,\,R,\,I) = \St(\rA_{n-1}, R, I)$ can be defined as the kernel $\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$.

For a general $I$ the relative Steinberg group $\St(\Phi, R, I)$ is no more a subgroup of $\St(\Phi, R)$ but rather is its central extension.
For the purposes of the present text it suffices to consider only relative Steinberg groups corresponding to splitting ideals $I$.
For more information regarding the general case we refer the reader to \cite[Section~3]{SCh}.

@@

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.
In the definition of the semidirect product the group $\St(n, R/I)$ acts on $\St(n, R, I)$ by conjugation via the splitting map, i.\,e. $g \cdot X(u,\,v) = X(\phi(s(g))u, \phi(s(g))^*v)$.

\begin{proof} For the sake of completeness we present below a more detailed version of Tulenbaev's proof.
Tulenbaev constructs a pair of mutually inverse maps
$$\xymatrix{ \St(n, R, I) \rtimes \St(n, R/I) \ar@<0.6ex>[r]^-\varphi &  \ar[l]^-\psi \St(n, R).}$$

Define the map $\psi$ by $\psi(x_{ij}(\xi)) = (X(e_i, (\xi - \pi(\xi)e_j),\ x_{ij}(\pi(\xi)))$. 
We have to check that $\psi$ preserves the relations \eqref{add0}--\eqref{ccf2}.
For example, let us check \eqref{ccf1}. 
First of all, recall that in any semidirect product of groups one can compute the commutator of two elements as follows:
$$[(a, b), (c, d)] = (a \cdot {}^bc \cdot {}^{bdb^{-1}}a^{-1} \cdot {}^{[b, d]}c^{-1} , [b, d]).$$
For $\xi\in R$ set $\xi' = \xi - \pi(\xi)$.
Specializing $a = X(e_i, \xi' e_j)$, $b = x_{ij}(\pi(\xi))$, $c = X(e_j, \eta'e_k)$, $d = x_{jk}(\pi(\eta))$ and using the definition of the conjugation action 
one computes the commutator $[\psi(x_{ij}(\xi), \psi(x_{jk}(\eta)]$ as follows:
$$\left(X(e_i, \xi' e_j) \cdot X(e_j + \pi(\xi)e_i, \eta'e_k) \cdot X(e_i, \pi(\eta)\xi' e_k - \xi' e_j) \cdot X(e_j, -\eta'e_k),\ x_{ik}(\pi(\xi\eta))\right).$$
After applying additivity relations \eqref{add2}, \eqref{add3} and moving the factor $X(e_i, -\xi' e_j)$ to the left hand side of the formula by means of relation \eqref{conj2}
one gets the following expression
$$(X(e_j, \eta'e_k) \cdot X(e_i, (\xi' \eta' + \pi(\xi)\eta' + \pi(\eta)\xi')e_k \cdot X(e_j, -\eta'e_k),\ x_{ik}(\pi(\xi\eta)).$$
Since $\xi'\eta' + \pi(\xi)\eta' + \pi(\eta)\xi' = \xi\eta - \pi(\xi\eta)$ the above expression clearly simplifies to $\psi(x_{ik}(\xi\eta))$, as claimed.
\end{proof}

\begin{rk} \label{rk:T123'} Notice that from the proof of the above proposition it follows that one could replace the family of relations \eqref{add3}
in the definition of $\St(n, R, I)$ with the following smaller family \eqref{add3'}.
\setcounter{equation}{2} \renewcommand{\theequation}{T\arabic{equation}'}
\begin{equation} X(e_ir+e_j,\,e_k)=X(e_i,\,rae_k)X(e_j,\,ae_k)\,\text{ for }r\in R,\ a\in I\text{ and } i,j,k\text{ distinct}. \label{add3'} \end{equation} \end{rk}

@@

\begin{df}
For $I\trianglelefteq R$ and $n\geq4$ define $\St^*(n,\,R,\,I)$ to be the group with the set of generators
\begin{multline*}
\{F(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}\,\cup\{S(u,\,v)\mid u\in I^n,\ v\in\E(n,\,R)e_1,\ u^tv=0\}
\end{multline*}
subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{R\arabic{equation}}
\begin{align}
&F(u,\,v)F(u,\,w)=F(u,\,v+w), \label{add4}\\
&S(u,\,v)S(w,\,v)=S(u+w,\,v), \label{add5}\\
&F(u,\,v)F(u',\,v')F(u,\,v)\inv=F(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj3} \\
&F(u,\,va)=S(ua,\,v),\ \text{for all}\ a\in I,\,(u,\,v^t)=(M e_1,\,e_2^t M\inv),\,M\!\in\E(n,\,R) \label{coinc}.
\end{align}
\end{df}
Notice that we only described how a generator $F(u, v)$ acts via conjugation on another generator $F(u',v')$ and have omitted three similar relations involving generators $S(u,v)$ of the second type.
The reason for this is the following lemma which asserts that the ``missing'' relations will follow automatically from \eqref{add4}--\eqref{coinc}.
\begin{lm}
\label{allyouneedisf}
Denote by $\phi\colon\St^*(n,\,R,\,I)\rightarrow\E(n,\,R)$ the natural map sending $F(u,\,v)\mapsto t(u,\,v)$ and $S(u,\,v)\mapsto t(u,\,v)$.
Then the following facts are true.
\begin{lmlist}
\item $\St^*(n,\,R,\,I)$ is generated as an abstract group by the set of elements 
      $$\{F(u,\,va)\mid a\in I,\ (u,\,v^t)=(Me_1,\,e_2^tM\inv),\,M\in\E(n,\,R)\};$$
\item for any $g\in\St^*(n,\,R,\,I)$ one has 
      \setcounter{equation}{2} \renewcommand{\theequation}{R\arabic{equation}'}
      \begin{equation} gF(u,\,v)g\inv=F(\phi(g)u,\,\phi(g\inv)^tv); \end{equation}
\item for any $g\in\St^*(n,\,R,\,I)$ one has
       \setcounter{equation}{2} \renewcommand{\theequation}{R\arabic{equation}''}
      \begin{equation} gS(u,\,v)g\inv=S(\phi(g)u,\,\phi(g\inv)^tv); \end{equation}
\item there is a ``transpose automorphism'' defined on $\St^*(n, R, I)$ satisfying
      $$F(u,\,v)^t=S(v,\,u),\quad S(u,\,v)^t=F(v,\,u).$$
\end{lmlist} \end{lm}

@@

In order to construct the arrow $\kappa$ we use the presentation from \cref{prop:TulPres}.
We set $$\kappa(X(u,v)) = F(u,v),\ u\in \E(n, R)e_1,\ v\in I^n.$$ 
In view of \cref{rk:T123'} it suffices to check that $F(u,v)$ satisfy the relations \eqref{add2}, \eqref{conj2}, \eqref{add3'}.
The validity relations \eqref{add2}, \eqref{conj2} is obvious while \eqref{add3'} follows from the following computation:
$$ F(e_i, rae_k) F(e_j, ae_k) = S(rae_i, e_k) S(ae_j, e_k) = S(a(re_i+e_j), e_k) = F(e_ir+e_j, ae_k). \qedhere $$

As for the map $\iota$ we firstly define it as a homomorphism to the absolute group (for which we de not need the assumption that $I$ is splitting).
$$\xymatrix{\St^*(n,\,R,\,I)\ar@<0.0ex>[r]^{\iota}&\St(n,\,R).}$$

Clearly, we must map the elements $F(u,\,v)$ to van der Kallen's generators $X(u,\,v)$.
We should find the images for the elements $S(u,\,v)$ as well.
First of all, recall that van der Kallen in~\cite[3.8--3.10]{vdK} constructs the following elements
$$x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.$$
The following definition is the ``transpose'' of \cite[3.13]{vdK}.
\begin{df} For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$
 consisiting of all elements $y\in\St(n,\,R)$ that admit the decomposition $y=\prod x(w^k,\,v)$, 
 where $\sum u^k=u$ and $u^k=c^k \cdot (e_pv_q-e_qv_p)$ for some $c^k\in R$, $1\leq p\neq q\leq n$. \end{df}

@@

@@

At this point we have demonstrated for $n\geq4$ and a splitting ideal $I$ that $\St^*(n,\,R,\,I)$ and $\St(n,\,R,\,I)$ are isomorphic.
Now we turn to the main result of this section, namely, we construct a map $$T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X],\,XR_a[X]),$$ for $n\geq4$. 
A local-global principle for Steinberg group and centrality of $\mathrm K_2$ formally follow from existance of this map~\cite[Lem.~15, Lem.~16, proof of Th.~2]{SCh}.
